---
title: "R Notebook"
output:
  html_notebook: default
  pdf_document: default
  html_document:
    df_print: paged
---

**Axon and Midbrain IP/Input Comparisons**

Load packages
```{r setup, results='hide'}
library(DESeq2)
library(tximport)
library(GenomicFeatures)
library(biomaRt)
library(reshape)
library(reshape2)
library(ggplot2)
library(pheatmap)
library(scater)
library(RColorBrewer)
library(dplyr)
library(tidyr)
library(tibble)
library(org.Mm.eg.db)
library(RUVSeq)
library(IHW)
library(Biobase)
library(preprocessCore)
library(EDASeq)
library(fgsea)
library(STRINGdb)
library(ggbeeswarm)
library(ggsci)
library(tidyverse)
library(EDASeq)
library(WGCNA)
library(rhdf5)
library(ggpubr)
source("SampleNetwork_1.06.r")


output_dir <- file.path('output',paste('deseq', Sys.Date(), format(Sys.time(), "%H:%M"), sep = '_'))
dir.create(output_dir, recursive = T)
base_dir <- getwd()
knitr::opts_knit$set(root.dir = output_dir)

sample_metadata_all <- read.csv(file.path(base_dir, 'sample_metadata_110719.csv'), header = T)
sample_metadata_all <- sample_metadata_all %>% filter(outlier == 'n' & ip_pool == 'n')
kallisto_path <- "/zfs/analysis/active_r_projects/trap_kall_protcod_ercc/"

load(file.path(base_dir, 'anno_mouse.RData'))
load(file.path(base_dir, 'tx2gene_mouse_ercc.RData'))

h5closeAll()
```

**Global parameters**
```{r}
alpha = 0.01
res_lfc = 0 #lfc threshold for DESeq2 results functions
lfc = 0 #lfc threshold for post-results filtering genes
replicate_replace = Inf
```

Load annotation database and mouse ERCC92 tx2gene
```{r}
# txdb <- makeTxDbFromGFF('/zfs/referenceGenomes/mouse/Mus_musculus.GRCm38.96.gtf')
# k <- keys(txdb, keytype = "TXNAME")
# tx2gene <- AnnotationDbi::select(txdb, k, "GENEID", "TXNAME")
# ERCC92 <- read.csv("ERCC92_names.txt", header = F)
# ERCC92$V2 <- ERCC92$V1
# colnames(ERCC92) <- c("TXNAME", "GENEID")
# tx2gene_ERCC <- bind_rows(tx2gene, ERCC92)
```

*Create functions for repeated code*
```{r}
run_deseq <- function() {
  #subset sample metadata
  sample_metadata <- sample_metadata_all %>% data.frame() %>%
                    as_tibble() %>% filter(eval(parse(text=metadata_filter)))
  assign(paste('sample_metadata', name, sep = '_'), envir = .GlobalEnv, sample_metadata)

  #create list of kallisto input files
  files <- file.path(kallisto_path, sample_metadata$sample_code, "abundance.h5")
  names(files) <- sample_metadata$sample_name
  
  #import kallisto files
  txi <- tximport(files, type = "kallisto", tx2gene = tx2gene_ERCC, ignoreTxVersion = T)
  
  #create deseq2 object
  dds <- DESeqDataSetFromTximport(txi, colData = sample_metadata, design = design)
  
  #filter low counts
  keep_feature <- rowMeans(counts(dds)) >= min_counts
  dds_removed <- dds[!keep_feature, ]
  dds <- dds[keep_feature, ]
  paste('Dimensions of sample data: ', dim(dds))
  
  #find ERCC spike-ins
  spikes <- rownames(dds)[grep("^ERCC", rownames(dds))]
  
  #calculate XYZ
  dds <- DESeq(dds, minReplicatesForReplace = replicate_replace)
  counts <- counts(dds, normalized = T)
  vsd <- DESeq2::vst(dds, blind = F)
  assign(paste('dds', name, sep = '_'), envir = .GlobalEnv, dds)
  assign(paste('dds_removed', name, sep = '_'), envir = .GlobalEnv, dds_removed)
  assign(paste('counts', name, sep = '_'), envir = .GlobalEnv, counts)
  assign(paste('vsd', name, sep = '_'), envir = .GlobalEnv, vsd)

  #summary(normalizationFactors(dds))
  
  #perform RUVseq
  eda_set <- newSeqExpressionSet(round(counts(dds, normalized = T)), phenoData = data.frame(sample_metadata, row.names=colnames(counts)))
  EDASeq::plotRLE(eda_set, outline = F, ylim = c(-3, 3), col = colors()[sample_metadata$ip])

  # eda_set1 <- betweenLaneNormalization(eda_set, which="full")
  # plotRLE(eda_set1, outline = F, ylim = c(-3, 3), col = colors()[sample_metadata$ip])
  
  # eda_set1 <- RUVg(eda_set, spikes, k=1)
  # pData(eda_set1)
  # plotRLE(eda_set1, outline = F, ylim = c(-3, 3), col = colors()[sample_metadata$ip])

  #create contrast variable
  contrast_name <- tail(resultsNames(dds), n=1)
  contrast_split <- strsplit(contrast_name, '_')
  contrast <- c(contrast_split[[1]][1], contrast_split[[1]][2], contrast_split[[1]][4])
  
  #calculate result
  res <- results(dds, alpha = alpha, lfcThreshold = res_lfc, contrast = contrast)
  assign(paste('res', name, sep = '_'), envir = .GlobalEnv, res)
  summary(res)
  
  #perform apeglm shrinkage
  res_shrunken <- lfcShrink(dds, coef = tail(resultsNames(dds), n=1), type = "apeglm", 
                            lfcThreshold = res_lfc, res = res, quiet = T)
  DESeq2::plotMA(res_shrunken, ylim=c(-5,5))
  assign(paste('res_shrunken', name, sep = '_'), envir = .GlobalEnv, res_shrunken)
  
  #add independent hypothesis weighting using basemean
  deRes <- as.data.frame(res)
  ihwRes <- ihw(pvalue ~ baseMean,  data = deRes, alpha = alpha)
  resIHW <- results(dds, filterFun=ihw, alpha = alpha, lfcThreshold = res_lfc, 
                    contrast = contrast)
  assign(paste('resIHW', name, sep = '_'), envir = .GlobalEnv, resIHW)
  
  res = resIHW
  
  #put results QC in (pvalue rejections, basemeans)

  #create results output object
  res_tb <- res %>%
    data.frame() %>%
    rownames_to_column(var="gene") %>% 
    as_tibble() %>%
    na.omit()

  sig <- res_tb %>%
    filter(padj < alpha & abs(log2FoldChange) > lfc)
  sig <- merge(as.data.frame(sig), anno, by.x = c("gene"), by.y = c("Gene.stable.ID"))
  sig <- merge(as.data.frame(sig), as.data.frame(counts(dds, normalized=TRUE)), 
               by.x = c("gene"), by.y = "row.names", sort=F)
  assign(paste('sig', name, sep = '_'), envir = .GlobalEnv, sig)
  
  res_tb <- merge(as.data.frame(res_tb), anno, by.x = c("gene"), by.y = c("Gene.stable.ID"))
  res_tb <- merge(as.data.frame(res_tb), as.data.frame(counts(dds, normalized=TRUE)), 
                  by.x = c("gene"), by.y = "row.names", sort=F)
  res_tb$padj[res_tb$padj == 0] <- .Machine$double.xmin #replacing absolute 0 values with smallest possible double for GSEA rank calculation
  assign(paste('res_output', name, sep = '_'), envir = .GlobalEnv, res_tb)
  
  res_tb$fcsign <- sign(res_tb$log2FoldChange)
  res_tb$logP <- -log10(res_tb$padj)
  res_tb$metric <- res_tb$logP/res_tb$fcsign
  res_tb <- res_tb %>% unique() %>% na.omit() %>% arrange(desc(metric))
  rank <- res_tb[,c("HGNC.symbol", "metric")]
  assign(paste('rank', name, sep = '_'), envir = .GlobalEnv, rank)
  
  write.csv(res_tb, file = paste('results', '_', name, '.csv', sep = ''))
  write.table(rank, file=paste('rank', '_', name, '.rnk', sep = ''), sep = '\t', quote=F, row.names=F)
}
```

**Comparison 1 -> Axon IP vs Midbrain IP**
```{r}
name = 'axon_ip_vs_mb_ip'

metadata_filter = "ip == 'ip' & batch_trap %in% c('b3')"

design = ~ collection_day + age + axon

min_counts = 100

run_deseq()
```

**Comparison 2 -> Axon UB vs Midbrain UB**
```{r}
name = 'axon_ub_vs_mb_ub'

metadata_filter = "ip == 'input' & batch_trap %in% c('b2','b3')"

design <- ~ batch_trap + age + axon

min_counts = 100

run_deseq()
```

**Comparison 3 -> Axon IP vs Axon UB**
```{r}
name = 'axon_ip_vs_axon_ub'

metadata_filter = "axon == 'axon'"

design <- ~ age + ip

min_counts = 10

run_deseq()
```

**Comparison 4 -> Midbrain IP vs Midbrain UB**
```{r}
name = 'mb_ip_vs_mb_ub'

metadata_filter = "axon == 'mb' & batch_trap %in% c('b2', 'b3')"

design <- ~ batch_trap + age + ip

min_counts = 10

run_deseq()
```

**Comparison 5 -> Dorsal Striatum IP vs Ventral Striatum IP**
```{r}
name = 'ds_ip_vs_vs_ip'

metadata_filter = "axon == 'axon' & ip == 'ip'"

design <- ~ collection_day + age + region

min_counts = 100

run_deseq()
```

**Comparison 6 -> Dorsal Striatum UB vs Ventral Striatum UB**
```{r}
name = 'ds_ub_vs_vs_ub'

metadata_filter = "axon == 'axon' & ip == 'input'"

design = ~ collection_day + age + region

min_counts = 100

run_deseq()
```

**Comparison 7 -> Dorsal Striatum IP vs Dorsal Striatum UB**
```{r}
name = 'ds_ip_vs_ds_ub'

metadata_filter = "region == 'ds'"

design = ~ age + ip

min_counts = 10

run_deseq()
```

**Comparison 8 -> Ventral Striatum IP vs Ventral Striatum UB**
```{r}
name = 'vs_ip_vs_vs_ub'

metadata_filter = "region == 'vs'"

design = ~ age + ip

min_counts = 10

run_deseq()
```

**Enrichment/Unique filtering**
```{r}
mb_enriched <- res_output_mb_ip_vs_mb_ub %>% filter(log2FoldChange > 0 & padj < 0.01)
mb_ip_genes <- nrow(mb_enriched)
axon_enriched <- res_output_axon_ip_vs_axon_ub %>% filter(log2FoldChange > 0 & padj < 0.01)
axon_ip_genes <- nrow(axon_enriched)

mb_axon_shared <- mb_enriched %>% filter(gene %in% axon_enriched$gene) %>% select(c("gene"))
mb_axon_genes <- nrow(mb_axon_shared)
mb_only <- mb_enriched %>% filter(!gene %in% axon_enriched$gene) %>% select(c("gene"))
axon_only <- axon_enriched %>% filter(!gene %in% mb_enriched$gene) %>% select(c("gene"))

ds_enriched <- res_output_ds_ip_vs_ds_ub %>% filter(log2FoldChange > 2 & padj < 0.001)
ds_ip_genes <- nrow(ds_enriched)
vs_enriched <- res_output_vs_ip_vs_vs_ub %>% filter(log2FoldChange > 2 & padj < 0.001)
vs_ip_genes <- nrow(vs_enriched)

ds_vs_shared <- vs_enriched %>% filter(gene %in% ds_enriched$gene) %>% select(c("gene"))
ds_only <- ds_enriched %>% filter(!gene %in% vs_enriched$gene) %>% select(c("gene"))
ds_only_genes <- nrow(ds_only)
vs_only <- vs_enriched %>% filter(!gene %in% ds_enriched$gene) %>% select(c("gene"))
vs_only_genes <- nrow(vs_only)

ds_ip_enriched_over_vs_ip <- res_output_ds_ip_vs_vs_ip %>% filter(log2FoldChange < 0.58 & padj < 0.001)
vs_ip_enriched_over_ds_ip <- res_output_ds_ip_vs_vs_ip %>% filter(log2FoldChange > 0.58 & padj < 0.001)
ds_ub_enriched_over_vs_ub <- res_output_ds_ub_vs_vs_ub %>% filter(log2FoldChange < 0.58 & padj < 0.001)
vs_ub_enriched_over_ds_ub <- res_output_ds_ub_vs_vs_ub %>% filter(log2FoldChange > 0.58 & padj < 0.001)
ds_ip_enriched_over_vs_ip <- ds_ip_enriched_over_vs_ip %>% filter(!gene %in% ds_ub_enriched_over_vs_ub$gene)
vs_ip_enriched_over_ds_ip <- vs_ip_enriched_over_ds_ip %>% filter(!gene %in% vs_ub_enriched_over_ds_ub$gene)

```

Plotting functions
```{r}
genes <- c('ENSMUSG00000030310', 'ENSMUSG00000030108', 'ENSMUSG00000030307', 'ENSMUSG00000021609') # GABA transporters
genes <- c('ENSMUSG00000033208', 'ENSMUSG00000030088', 'ENSMUSG00000020932', 'ENSMUSG00000053279', 'ENSMUSG00000005360', 'ENSMUSG00000026473', 'ENSMUSG00000005089', 'ENSMUSG00000000214', 'ENSMUSG00000021609') # Astrocyte markers

all_counts <- as.data.frame(log2(counts(dds_axon_ip_vs_axon_ub, normalized = T)))

counts_to_plot <- all_counts %>% 
  rownames_to_column(var = "gene") %>% 
  as_tibble %>%
  filter(gene %in% genes) %>% 
  gather(sample,counts,-gene) %>% 
  extract(sample, c('ip', 'region', 'age', 'day'), regex = "([A-Z]{2})([A-Z]{2})([0-9]*)(.*)", remove = TRUE, convert = FALSE) %>%
  inner_join(anno, by = c("gene" = "Gene.stable.ID"))
  
# display.brewer.all(colorblindFriendly = TRUE)

ip_ub_multi_violinplot <- ggplot(counts_to_plot, aes(factor(HGNC.symbol), counts, fill = factor(ip))) + 
  geom_violin() + 
  geom_quasirandom(alpha = 0.4, dodge.width=1, size = 0.75) + 
  theme_bw() + 
  labs(fill='Sample Type') + 
  xlab("Gene") + 
  ylab("log2 Counts") +
  scale_color_npg() +
  scale_fill_npg()

ip_ub_multi_violinplot

ip_ub_multi_boxplot <- ggplot(counts_to_plot, aes(x = HGNC.symbol, y = counts, fill = ip)) +
  geom_boxplot(outlier.size = 0) +
  geom_point(pch = 21, position = position_jitterdodge()) +
  theme_bw() + 
  labs(fill = 'Sample Type') +
  xlab('Gene') +
  ylab('log2 Counts') +
  scale_color_npg() +
  scale_fill_npg() +
  ggtitle('Astrocyte markers in Axon TRAP samples')

ip_ub_multi_boxplot

#+ facet_grid(age ~ region)

```

More plotting functions
```{r}
genes <- c('ENSMUSG00000021609')

ip_ub_single_violinplot <- ggplot(counts_to_plot, aes(factor(ip), counts)) + 
  geom_violin(aes(fill = factor(ip))) + 
  geom_jitter(height = 0, width = 0.1) + 
  labs(fill='Sample') + 
  xlab('Sample') + 
  ylab("log2 Counts") + 
  theme_bw() + 
  scale_color_npg() +
  scale_fill_npg()

ip_ub_single_violinplot

ip_ub_single_boxplot <- ggplot(counts_to_plot, aes(factor(ip), counts)) + 
  geom_boxplot(aes(fill = factor(ip))) + 
  geom_jitter(height = 0, width = 0.1) + 
  labs(fill='Sample') + 
  xlab(genes) + 
  ylab("log2 Counts") + 
  theme_bw() + 
  scale_color_npg() +
  scale_fill_npg()

ip_ub_single_boxplot

```

ER/Glycosylation/Oligosaccharide plots
```{r}
er_chaperone_genes <- c('ENSMUSG00000020048', 'ENSMUSG00000026864', 'ENSMUSG00000032115', 'ENSMUSG00000024357', 'ENSMUSG00000004460', 'ENSMUSG00000027006', 'ENSMUSG00000022136', 'ENSMUSG00000019802', 'ENSMUSG00000026740', 'ENSMUSG00000070436', 'ENSMUSG00000020048', 'ENSMUSG00000023973')

glycosylation_genes <- c('ENSMUSG00000020368', 'ENSMUSG00000003814', 'ENSMUSG00000089960', 'ENSMUSG00000030062', 'ENSMUSG00000027642', 'ENSMUSG00000071650', 'ENSMUSG00000040462', 'ENSMUSG00000020311', 'ENSMUSG00000036646', 'ENSMUSG00000038312', 'ENSMUSG00000043019')

oligosaccharide_genes <- c('ENSMUSG00000039427', 'ENSMUSG00000063362', 'ENSMUSG00000033809', 'ENSMUSG00000039740', 'ENSMUSG00000052395', 'ENSMUSG00000032059', 'ENSMUSG00000035845', 'ENSMUSG00000073792', 'ENSMUSG00000035704', 'ENSMUSG00000032116', 'ENSMUSG00000032437', 'ENSMUSG00000030035', 'ENSMUSG00000038803', 'ENSMUSG00000036632', 'ENSMUSG00000078919')

genes <- er_chaperone_genes

all_results <- res_output_axon_ip_vs_axon_ub %>% inner_join(res_output_mb_ip_vs_mb_ub, by = c("gene" = "gene"))

results_to_plot <- all_results %>% 
  as_tibble %>%
  filter(gene %in% genes) %>% 
  select(Gene.name.x, log2FoldChange.x, log2FoldChange.y) %>% 
  rename(gene = Gene.name.x, axon = log2FoldChange.x, soma = log2FoldChange.y) %>% 
  gather(region, enrichment, -gene)

ggplot(results_to_plot, aes(x = gene, y = enrichment, fill = region)) + 
  geom_bar(stat = "identity", position = "dodge", colour="black") +
  theme_bw() +
  coord_flip() +
  scale_color_npg() +
  scale_fill_npg() +
  ggtitle('Glycosylation Gene IP Expression Relative to Input') +
  labs(fill='Region') + 
  xlab('Gene') + 
  ylab('Log2 fold change compared to input') +
  geom_hline(yintercept = 0)

```

Normalisation factor plots
```{r}
nf <- as.data.frame(normalizationFactors(dds_axon_ip_vs_axon_ub)) 
nf <- nf %>%
  rownames_to_column(var = 'gene') %>%
  as_tibble() %>%
  gather(colnames(nf), key = 'samplename', value = 'normFactors') %>%
  separate(samplename, c('ip', 'region', 'sampleinfo'), sep = c(2, 4))

ggplot(nf, aes(normFactors, group = ip, fill = ip)) + 
  geom_density(alpha = 0.8, position = 'identity') + 
  theme_bw() +
  xlim(0.5, 2) +
  scale_color_npg() +
  scale_fill_npg() +
  ggtitle('Normalisation Factor Distribution') +
  labs(fill='Sample Type') + 
  xlab('Normalisation Factor') + 
  ylab('Density') +
  geom_hline(yintercept = 0)

#ggsave("nf_plot_mb_batch_2.pdf", width = 20, height = 20, units = "cm")
```

Cre/eGFP unbound plots
```{r}
trinity_tx2gene <- read.csv("/zfs/analysis/PK_TRAP/trinity_out_dir/tx_names.txt", header = F) %>%
  rename(TXID = V1) %>%
  mutate(GENEID = TXID) %>%
  mutate(GENEID = recode(GENEID, TRINITY_DN144_c0_g1_i8 = "eGFP-L10a", TRINITY_DN144_c0_g5_i1 = "IRES-Cre")) %>%
  as_tibble()

trinity_ub_path <- "/zfs/analysis/PK_TRAP/trinity_kallisto_all/"

name = 'axon_ub_vs_mb_ub'
metadata_filter = "ip == 'input' & batch_trap %in% c('b2','b3')"
design <- ~ batch_trap + age + axon

trinity_ub_sample_metadata <- sample_metadata_all %>% data.frame() %>%
                    as_tibble() %>% filter(eval(parse(text=metadata_filter)))

#create list of kallisto input files
trinity_ub_files <- file.path(trinity_ub_path, trinity_ub_sample_metadata$sample_code, "abundance.h5")
names(trinity_ub_files) <- trinity_ub_sample_metadata$sample_name

#import kallisto files
trinity_ub_txi <- tximport(trinity_ub_files, type = "kallisto", tx2gene = trinity_tx2gene, ignoreTxVersion = F, geneIdCol = "GENEID")
  
#create deseq2 object
trinity_ub_dds <- DESeqDataSetFromTximport(trinity_ub_txi, colData = trinity_ub_sample_metadata, design = design)

#calculate XYZ
trinity_ub_dds <- DESeq(trinity_ub_dds, minReplicatesForReplace = replicate_replace)

trinity_ub_counts <- as.data.frame(counts(trinity_ub_dds, normalized = T)) %>%
  rownames_to_column(var = "gene") %>%
  as_tibble() %>%
  gather(sample,counts,-gene) %>%
  inner_join(sample_metadata_all, by = c('sample' = 'sample_name')) 

  egfp_l10a_ub_counts <- trinity_ub_counts %>%
  filter(gene == 'eGFP-L10a') %>%
  select(gene, sample, region, ip, counts) 

ires_cre_ub_counts <- trinity_ub_counts %>%
  filter(gene == 'IRES-Cre') %>%
  select(gene, sample, region, ip, counts) 

#eGFP Input samples
caption <- 'eGFP is detectable in input RNA from striatum, either supporting axonal translation of eGFP-L10a or transcription/translation in non-dopamine cell types.'
caption <- paste0(strwrap(caption, 120), sep="", collapse="\n")

eGFP_multi_boxplot <- ggplot(egfp_l10a_ub_counts, aes(x = region, y = counts, fill = region)) +
  geom_boxplot(outlier.size = 0) +
  geom_jitter(alpha = 0.5, pch = 21) +  
  theme_bw() + 
  labs(fill = 'Region') +
  xlab('Region') +
  ylab('Counts') +
  expand_limits(y = 0) +
  scale_color_npg() +
  scale_fill_npg() +
  labs(fill = 'Region',
       title = 'eGFP-L10a Transcript Counts in Input RNA', 
       caption = caption) +
  theme(plot.caption = element_text(hjust = 0, face = "bold"), legend.position = "none")

eGFP_multi_boxplot

#Cre Input samples
caption <- 'Cre is not detectable in input RNA from striatum, however counts are low, so this does not rule out Cre expression in striatum.'
caption <- paste0(strwrap(caption, 120), sep="", collapse="\n")

cre_multi_boxplot <- ggplot(ires_cre_ub_counts, aes(x = region, y = counts, fill = region)) +
  geom_boxplot(outlier.size = 0) +
  geom_jitter(alpha = 0.5, pch = 21) +
  theme_bw() + 
  labs(fill = 'Region') +
  xlab('Region') +
  ylab('Counts') +
  expand_limits(y = 0) +
  scale_color_npg() +
  scale_fill_npg() +
  labs(fill = 'Region',
       title = 'IRES-Cre Transcript Counts in Input RNA', 
       caption = caption) +
  theme(plot.caption = element_text(hjust = 0, face = "bold"), legend.position = "none")

cre_multi_boxplot

```

Cre/eGFP IP plots
```{r}
trinity_ip_path <- "/zfs/analysis/PK_TRAP/trinity_kallisto_all/"

name = 'axon_ip_vs_mb_ip'
metadata_filter = "ip == 'ip' & batch_trap %in% c('b2', 'b3')"
design <- ~ age + axon

trinity_ip_sample_metadata <- sample_metadata_all %>% data.frame() %>%
                    as_tibble() %>% filter(eval(parse(text=metadata_filter)))

#create list of kallisto input files
trinity_ip_files <- file.path(trinity_ip_path, trinity_ip_sample_metadata$sample_code, "abundance.h5")
names(trinity_ip_files) <- trinity_ip_sample_metadata$sample_name

#import kallisto files
trinity_ip_txi <- tximport(trinity_ip_files, type = "kallisto", tx2gene = trinity_tx2gene, ignoreTxVersion = F, geneIdCol = "GENEID")
  
#create deseq2 object
trinity_ip_dds <- DESeqDataSetFromTximport(trinity_ip_txi, colData = trinity_ip_sample_metadata, design = design)

#calculate XYZ
trinity_ip_dds <- DESeq(trinity_ip_dds, minReplicatesForReplace = replicate_replace)

trinity_ip_counts <- as.data.frame(counts(trinity_ip_dds, normalized = T)) %>%
  rownames_to_column(var = "gene") %>%
  as_tibble() %>%
  gather(sample,counts,-gene) %>%
  inner_join(sample_metadata_all, by = c('sample' = 'sample_name')) 

ires_cre_ip_counts <- trinity_ip_counts %>%
  filter(gene == 'IRES-Cre') %>%
  select(gene, sample, region, ip, counts) 

#Cre IP samples
caption <- 'Cre is below the limit of detection in axonal IP samples. It is either present at a low level in non-dopaminergic cells or not transcribed at all. qPCR needs to be done on striatal input samples to accurately compare expression to midbrain.'
caption <- paste0(strwrap(caption, 120), sep="", collapse="\n")

cre_multi_boxplot <- ggplot(ires_cre_ip_counts, aes(x = region, y = counts, fill = region)) +
  geom_boxplot(outlier.size = 0) +
  geom_jitter(alpha = 0.5, pch = 21) +
  theme_bw() + 
  labs(fill = 'Region') +
  xlab('Region') +
  ylab('Counts') +
  expand_limits(y = 0) +
  scale_color_npg() +
  scale_fill_npg() +
  labs(fill = 'Region',
       title = 'IRES-Cre Transcript Counts in IP RNA', 
       caption = caption) +
  theme(plot.caption = element_text(hjust = 0, face = "bold"), legend.position = "none")

cre_multi_boxplot

```

Cre/eGFP IP vs UB plots
```{r}
trinity_ip_ub_path <- "/zfs/analysis/PK_TRAP/trinity_kallisto_all/"

name = 'ip_vs_ub'
metadata_filter = "batch_trap %in% c('b2', 'b3')"
design <- ~ age + ip

trinity_ip_ub_sample_metadata <- sample_metadata_all %>% data.frame() %>%
                    as_tibble() %>% filter(eval(parse(text=metadata_filter)))

#create list of kallisto input files
trinity_ip_ub_files <- file.path(trinity_ip_ub_path, trinity_ip_ub_sample_metadata$sample_code, "abundance.h5")
names(trinity_ip_ub_files) <- trinity_ip_ub_sample_metadata$sample_name

#import kallisto files
trinity_ip_ub_txi <- tximport(trinity_ip_ub_files, type = "kallisto", tx2gene = trinity_tx2gene, ignoreTxVersion = F, geneIdCol = "GENEID")
  
#create deseq2 object
trinity_ip_ub_dds <- DESeqDataSetFromTximport(trinity_ip_ub_txi, colData = trinity_ip_ub_sample_metadata, design = design)

#calculate XYZ
trinity_ip_ub_dds <- DESeq(trinity_ip_ub_dds, minReplicatesForReplace = replicate_replace)
trinity_ip_ub_counts <- as.data.frame(counts(trinity_ip_ub_dds, normalized = T)) %>%
  rownames_to_column(var = "gene") %>%
  as_tibble() %>%
  gather(sample,counts,-gene) %>%
  inner_join(sample_metadata_all, by = c('sample' = 'sample_name')) 

  egfp_l10a_ip_ub_counts <- trinity_ip_ub_counts %>%
  filter(gene == 'eGFP-L10a') %>%
  select(gene, sample, region, ip, counts) 

ires_cre_ip_ub_counts <- trinity_ip_ub_counts %>%
  filter(gene == 'IRES-Cre') %>%
  select(gene, sample, region, ip, counts) 

#eGFP IP and UB samples
caption <- 'eGFP-L10a is enriched in midbrain samples, but not in axonal samples. This suggests that eGFP-L10a RNA is not being translated in striatum, while Th and Slc6a3 are.'
caption <- paste0(strwrap(caption, 100), sep="", collapse="\n")

eGFP_multi_boxplot <- ggplot(egfp_l10a_ip_ub_counts, aes(x = region, y = counts, fill = ip)) +
  geom_boxplot(outlier.size = 0) +
  geom_point(pch = 21, position = position_jitterdodge()) +
  theme_bw() + 
  xlab('Gene') +
  ylab('Counts') +
  expand_limits(y = 0) +
  scale_color_npg() +
  scale_fill_npg() +
  labs(fill = 'Region',
       title = 'eGFP-L10a Transcript Counts in IP and Input RNA', 
       caption = caption) +
  theme(plot.caption = element_text(hjust = 0, face = "bold"), legend.position = "right")

eGFP_multi_boxplot

#Cre IP and UB samples
caption <- 'Cre transcripts are enriched in midbrain and axon samples, however qPCR needs to be done to accurately measure axonal enrichment because of low count number.'
caption <- paste0(strwrap(caption, 100), sep="", collapse="\n")

cre_multi_boxplot <- ggplot(ires_cre_ip_ub_counts, aes(x = region, y = counts, fill = ip)) +
  geom_boxplot(outlier.size = 0) +
  geom_point(pch = 21, position = position_jitterdodge()) +
  theme_bw() + 
  labs(fill = 'Region') +
  xlab('Region') +
  ylab('Counts') +
  expand_limits(y = 0) +
  scale_color_npg() +
  scale_fill_npg() +
  labs(fill = 'Sample Type',
       title = 'IRES-Cre Transcript Counts in IP and Input RNA', 
       caption = caption) +
  theme(plot.caption = element_text(hjust = 0, face = "bold"), legend.position = "right")

cre_multi_boxplot

```

Using striatal single cell markers to examine cell type enrichment in striatum [Link to paper](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5004635/)
Used Table S2 to generate list of matching genes
```{r}
striatal_markers <- read.csv(file = "/zfs/analysis/active_r_projects/pk_trap/striatal_markers.csv")

res_output_axon_ip_vs_axon_ub %>% 
  filter(Gene.name %in% striatal_markers$gene) %>%
  inner_join(striatal_markers, by = c('Gene.name' = 'gene')) %>%
  filter(padj < 0.01) %>%
  select(Gene.name, Cell.Type, log2FoldChange) %>%
  arrange(desc(log2FoldChange))

```

Evaluate astrocyte enrichment in [Sakers, 2017 dataset](https://www.pnas.org/content/114/19/E3830)
```{r}
kallisto_path <- "/zfs/analysis/PK_TRAP/astrocyte_trap_sakers_2017/kallisto/"
sample_metadata_all <- read.csv("/zfs/analysis/active_r_projects/pk_trap/astro_sample_table.csv")
name = 'sakers_ip_vs_ub'
metadata_filter = "ip %in% c('ip', 'ub')"
design <- ~ ip
reference_level <- 'ub'
min_counts = 10

run_deseq()

```
Midbrain WGCNA
```{r}
name = 'mb_18m_wt_vs_ovx'

metadata_filter = "region == 'mb' & ip == 'ip' & age == 'old' & batch_trap == 'b2' & ip_pool == 'n'"

design = ~ sex + collection_day + ovx

min_counts = 200

run_deseq()

plotPCA(vsd_mb_18m_wt_vs_ovx, intgroup = 'ovx')

#Checking for good genes
gsg = goodSamplesGenes(assay(dds_mb_18m_wt_vs_ovx), verbose = 5);
gsg$allOK

vsd_mb_18m_wt_vs_ovx_corrected <- vst(dds_mb_18m_wt_vs_ovx, blind = F)
mat <- assay(vsd_mb_18m_wt_vs_ovx_corrected)
mat <- limma::removeBatchEffect(mat, vsd_mb_18m_wt_vs_ovx_corrected$collection_day)
assay(vsd_mb_18m_wt_vs_ovx_corrected) <- mat
datExprT <- assay(vsd_mb_18m_wt_vs_ovx_corrected)
datExprT <- as.data.frame(datExprT) %>%
  rownames_to_column(var = "gene")

all(sample_metadata_mb_18m_wt_vs_ovx$sample_name == colnames(datExprT[2:17]))

#datExprT <- counts_mb_18m_wt_vs_ovx
sample_metadata_mb_18m_wt_vs_ovx <- as.data.frame(sample_metadata_mb_18m_wt_vs_ovx[-c(3,4)])

SampleNetwork(
  datExprT=datExprT,
  method1="correlation",
  impute1=FALSE,
  subset1=NULL,
  skip1=1,
  indices1=list(c(colnames(datExprT[2:17]))),
  sampleinfo1=sample_metadata_mb_18m_wt_vs_ovx,
  subgroup1=13,
  samplelabels1=1,
  grouplabels1=5,
  fitmodels1=TRUE,
  whichmodel1="multivariate",
  whichfit1="pc1",
  btrait1=c(3,6,13),
  trait1=c(13),
  asfactors1=c(3,6,13),
  projectname1="mb_18m_wt_vs_ovx",
  cexlabels1=0.7,
  normalize1=TRUE,
  replacenegs1=FALSE,
  exportfigures1=TRUE,
  verbose=TRUE
)

input = as.data.frame(t(datExprT[, -c(1)]))
names(input)<- datExprT$gene
dim(input)

#WGCNA Analysis
powers = c(c(1:11), seq(from = 12, to=20, by=2))
sft = pickSoftThreshold(input,corFnc = "bicor", corOptions=list(maxPOutliers=0.1), networkType = "signed hybrid", powerVector = powers, verbose = 5)

# Plot the results:
par(mfrow = c(1,2));
cex1 = 0.9;
# Scale-free topology fit index as a function of the soft-thresholding power (Signed)
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
     main = paste("Scale independence"));
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     labels=powers,cex=cex1,col="red");
# this line corresponds to using an R^2 cut-off of h
abline(h=0.90,col="red")
# Mean connectivity as a function of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5],
     xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
     main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")

sft

power <- 9
  
##AUTOMATIC METHOD!
  
datExpr <- input
datTraits <- traits

traits <- sample_metadata_mb_18m_wt_vs_ovx[c(13)]
traits$ovx <- as.numeric(traits$ovx)-1
rownames(traits) <- rownames(input)
  
net = blockwiseModules(datExpr, power = power,maxBlockSize = 10000,
                             TOMType = "signed Nowick", minModuleSize = 30, deepSplit = 2,
                             reassignThreshold = 0, mergeCutHeight = 0.25,
                             numericLabels = TRUE, pamRespectsDendro = FALSE,
                             saveTOMs = FALSE, networkType = "signed hybrid",
                             verbose = 5,corType = "bicor", maxPOutliers = 0.1,
                             pearsonFallback = "individual")

# Convert labels to colors for plotting
mergedColors = labels2colors(net$colors)
# Plot the dendrogram and the module colors underneath
plotDendroAndColors(net$dendrograms[[1]], mergedColors[net$blockGenes[[1]]],
                  "Module colors",
                  dendroLabels = FALSE, hang = 0.03,
                  addGuide = TRUE, guideHang = 0.05)

moduleLabels = net$colors
moduleColors = labels2colors(net$colors)
MEs = net$MEs;
geneTree = net$dendrograms[[1]];

# Define numbers of genes and samples
nGenes = ncol(datExpr);
nSamples = nrow(datExpr);
# Recalculate MEs with color labels
MEs0 = moduleEigengenes(datExpr, moduleColors)$eigengenes
MEs = orderMEs(MEs0)
moduleTraitCor = cor(MEs, datTraits, use = "p");
moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples);


moduleTraitCorInteresting <- as.data.frame(moduleTraitCor) %>%
  rownames_to_column("module") %>%
  filter(ovx > 0.5 | ovx  < -0.5) %>%
  column_to_rownames("module") %>%
  as.matrix()

moduleTraitPvalueInteresting <- as.data.frame(moduleTraitPvalue) %>%
  rownames_to_column("module") %>%
  filter(ovx < 0.1 & module %in% rownames(moduleTraitCorInteresting)) %>%
  column_to_rownames("module") %>%
  as.matrix()

# moduleTraitInteresting <- moduleTraitCorInteresting %>%
#   inner_join(y=moduleTraitPvalueInteresting, by = "module", suffix = c(".cor", ".pval")) %>%
#   column_to_rownames("module") %>%
#   as.matrix()


# Will display correlations and their p-values
textMatrix = paste(signif(moduleTraitCorInteresting, 2), "\n(",
                        signif(moduleTraitPvalueInteresting, 1), ")", sep = "");
dim(textMatrix) = dim(moduleTraitCorInteresting)

# Display the correlation values within a heatmap plot
labeledHeatmap(Matrix = moduleTraitCorInteresting,
             xLabels = names(datTraits),
             yLabels = rownames(moduleTraitCorInteresting),
             ySymbols = rownames(moduleTraitCorInteresting),
             colorLabels = FALSE,
             colors = greenWhiteRed(50),
             textMatrix = textMatrix,
             setStdMargins = FALSE,
             cex.text = 0.5,
             zlim = c(-1,1),
             main = paste("Module-trait relationships"))

# Define variable ovx containing the ovx column of datTrait
ovx = as.data.frame(datTraits$ovx);
names(ovx) = "ovx"
# names (colors) of the modules
modNames = substring(names(MEs), 3)
geneModuleMembership = as.data.frame(cor(datExpr, MEs, use = "p"));
MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples));
names(geneModuleMembership) = paste("MM", modNames, sep="");
names(MMPvalue) = paste("p.MM", modNames, sep="");
geneTraitSignificance = as.data.frame(cor(datExpr, ovx, use = "p"));
GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples));
names(geneTraitSignificance) = paste("GS.", names(ovx), sep="");
names(GSPvalue) = paste("p.GS.", names(ovx), sep="");

module = "blue"
column = match(module, modNames);
moduleGenes = moduleColors==module;

verboseScatterplot(abs(geneModuleMembership[moduleGenes, column]),
                   abs(geneTraitSignificance[moduleGenes, 1]),
                 xlab = paste("Module Membership in", module, "module"),
                 ylab = "Gene significance for OVX status",
                 main = paste("Module membership vs. gene significance\n"),
                 cex.main = 1.2, cex.lab = 1.2, cex.axis = 1.2, col = module)

blue_module_mm_sig <-  tibble(module_membership = abs(geneModuleMembership[moduleGenes, column]), trait_significance = abs(geneTraitSignificance[moduleGenes, 1]))

blue_module_plot <- ggplot(blue_module_mm_sig, aes(x = module_membership, y = trait_significance)) + 
  geom_point() + 
  theme_classic() + 
  xlab('Module Membership') +
  ylab('Correlation with OVX Status') +
  expand_limits(y = 0) +
  scale_color_npg() +
  scale_fill_npg() +
  labs(fill = 'Region',
       title = 'Correlation of Module Membership with OVX Status',
       caption = "Genes situated in the top right of the graph are most likely to be of interest.") +
  theme(plot.title = element_text(hjust = 0, face = "bold"), plot.caption = element_text(hjust = 0, face = "bold"))

blue_module_plot + stat_cor(method = "pearson")

#write.csv(file = "ovx_module_1.csv", names(datExpr)[moduleColors=="brown"])

```

```{r}
sampleTree = hclust(dist(input), method = "average");
par(cex = 0.6);
par(mar = c(0,4,2,0))
plot(sampleTree, main = "Sample clustering to detect outliers", sub="", xlab="", cex.lab = 1.5,
     cex.axis = 1.5, cex.main = 2)

n_genes = ncol(input)
n_samples = nrow(input)
  
traits <- sample_metadata_mb_18m_wt_vs_ovx[c(3, 6, 13)]
traits$collection_day <- as.numeric(traits$collection_day)-1
traits$sex <- as.numeric(traits$sex)-1
traits$ovx <- as.numeric(traits$ovx)-1
rownames(traits) <- rownames(input)

# Re-cluster samples
sampleTree2 = hclust(dist(input), method = "average")
# Convert traits to a color representation: white means low, red means high, grey means missing entry
traitColors = numbers2colors(traits, signed = T);
# Plot the sample dendrogram and the colors underneath.
plotDendroAndColors(sampleTree2, traitColors,
                    groupLabels = names(traits),
                    main = "Sample dendrogram and trait heatmap")
  
adjacency = adjacency(input, power = power, type = "signed", corFnc = "bicor", corOptions = list(maxPOutliers = 0.05, pearsonFallback = "individual"));
  
TOM = TOMsimilarity(adjacency, TOMType = "signed", verbose = 3);
  
dissTOM = 1-TOM

nSelect = 900
# For reproducibility, we set the random seed
set.seed(10);
select = sample(n_genes, size = nSelect);
selectTOM = dissTOM[select, select];
# There’s no simple way of restricting a clustering tree to a subset of genes, so we must re-cluster.
selectTree = hclust(as.dist(selectTOM), method = "average")
plotDiss = selectTOM^7;
diag(plotDiss) = NA;
TOMplot(plotDiss, selectTree, main = "Topological Overlap Matrix of 18 month IP samples")
  
geneTree = hclust(as.dist(dissTOM), method = "average");
plot(geneTree, xlab="", sub="", main = "Gene clustering on TOM-based dissimilarity",
       labels = FALSE, hang = 0.04);

CUT_HEIGHT <- 0.995

abline(h = CUT_HEIGHT, col = "red");
clust = cutreeStatic(sampleTree, cutHeight = CUT_HEIGHT, minSize = 10)

# We like large modules, so we set the minimum module size relatively high:
minModuleSize = 30;
# Module identification using dynamic tree cut:
dynamicMods = cutreeDynamic(dendro = geneTree, distM = dissTOM,
                            deepSplit = 2, pamRespectsDendro = FALSE,
                            minClusterSize = minModuleSize,
                            method = "hybrid");
table(dynamicMods)

# Convert numeric lables into colors
dynamicColors = labels2colors(dynamicMods)
table(dynamicColors)
# Plot the dendrogram and colors underneath
plotDendroAndColors(geneTree, dynamicColors, "Dynamic Tree Cut",
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05, 
                    main = "Gene dendrogram and module colors")
  
  # Calculate eigengenes
  MEList = moduleEigengenes(input, colors = dynamicColors)
  MEs = MEList$eigengenes
  # Calculate dissimilarity of module eigengenes
  MEDiss = 1-cor(MEs);
  # Cluster module eigengenes
  METree = hclust(as.dist(MEDiss), method = "average");
  # Plot the result
  plot(METree, main = "Clustering of module eigengenes",
       xlab = "", sub = "")
  
  MEDissThres = 0.25
  # Plot the cut line into the dendrogram
  abline(h=MEDissThres, col = "red")
  # Call an automatic merging function
  merge = mergeCloseModules(input, dynamicColors, cutHeight = MEDissThres, verbose = 3)
  # The merged module colors
  mergedColors = merge$colors;
  # Eigengenes of the new merged modules:
  mergedMEs = merge$newMEs;
  
  plotDendroAndColors(geneTree, cbind(dynamicColors, mergedColors),
                      c("Dynamic Tree Cut", "Merged dynamic"),
                      dendroLabels = FALSE, hang = 0.03,
                      addGuide = TRUE, guideHang = 0.05)
  
  moduleColors = mergedColors
  # Construct numerical labels corresponding to the colors
  colorOrder = c("grey", standardColors(50));
  moduleLabels = match(moduleColors, colorOrder)-1;
  MEs = mergedMEs;
  
    #Calculate module eigengenes and correlate eigengenes to trait
  MEs0 = moduleEigengenes(input, moduleColors)$eigengenes
  MEs = orderMEs(MEs0)
  moduleTraitCor = cor(MEs, traits, use = "p");
  moduleTraitPvalue = corPvalueStudent(moduleTraitCor, n_samples);
  
  # Will display correlations and their p-values
  textMatrix = paste(signif(moduleTraitCor, 2), "\n(",
                     signif(moduleTraitPvalue, 1), ")", sep = "");
  dim(textMatrix) = dim(moduleTraitCor)
  par(mar = c(6, 8.5, 3, 3));
  # Display the correlation values within a heatmap plot
  labeledHeatmap(Matrix = moduleTraitCor,
                 xLabels = names(traits),
                 yLabels = names(MEs),
                 ySymbols = names(MEs),
                 colorLabels = FALSE,
                 colors = greenWhiteRed(50),
                 textMatrix = textMatrix,
                 setStdMargins = FALSE,
                 cex.text = 0.5,
                 zlim = c(-1,1),
                 main = paste("Module-trait relationships"))

moduleTraitCor1 <- as.data.frame(moduleTraitCor) %>%
  rownames_to_column("module") %>%
  filter(ovx > 0.5 | ovx < -0.5)

moduleTraitPvalue1 <- as.data.frame(moduleTraitPvalue) %>%
  rownames_to_column("module") %>%
  filter(ovx < 0.1)

  # Define variable X containing the X column of traits
  ovx = as.data.frame(traits$ovx);
  names(ovx) = "ovx"
  modNames = substring(names(MEs), 3)
  geneModuleMembership = as.data.frame(cor(input, MEs, use = "p"));
  MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), n_samples));
  names(geneModuleMembership) = paste("MM", modNames, sep="");
  names(MMPvalue) = paste("p.MM", modNames, sep="");
  geneTraitSignificance = as.data.frame(cor(input, ovx, use = "p"));
  GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), n_samples));
  names(geneTraitSignificance) = paste("GS.", names(ovx), sep="");
  names(GSPvalue) = paste("p.GS.", names(ovx), sep="");

  module = "coral"
  column = match(module, modNames);
  moduleGenes = moduleColors==module;
  sizeGrWindow(7, 7);
  par(mfrow = c(1,1));
  verboseScatterplot(abs(geneModuleMembership[moduleGenes, column]),
                     abs(geneTraitSignificance[moduleGenes, 1]),
                     xlab = paste("Module Membership in", module, "module"),
                     ylab = "Gene significance for disease",
                     main = paste("Module membership vs. gene significance\n"),
                     cex.main = 1.2, cex.lab = 1.2, cex.axis = 1.2, col = "blue")


  which.module="coral";
  plotMat(input[,colorh1==which.module], nrgcols=30,
          clabels=rownames(input),rcols=which.module)
  
  datME=moduleEigengenes(heatmap_INPUT,colorh1)$eigengenes
  
  sizeGrWindow(8,7);
  which.module="brown"
  ME=datME[, paste("ME",which.module, sep="")]
  par(mfrow=c(2,1), mar=c(0.3, 5.5, 4, 2))
  plotMat(t(scale(heatmap_INPUT[,colorh1==which.module ]) ),
          nrgcols=30,rlabels=F,rcols=which.module, cex.main=1, clabels=rownames(heatmap_INPUT))
  par(mar=c(5, 4.2, 0, 0.7))
  barplot(ME, col=which.module, main="", cex.main=2,
          ylab="eigengene expression")

```

